Emitter Source Geolocation from Imparted Rotor Blade Modulation

### 2016-11-26 Background.tex
remove red markers, get ready to publish

\section{Thesis Scope}
\textcolor{red}{EDITED}

The scope of this thesis is to investigate whether it is possible to use the effect of RBM to geolocate a transmitter in ideal conditions. Under ideal conditions the received signal is deterministic based upon the propagation model and the initial conditions. This thesis does not intend to include every physical effect but is limited to the effects defined in the propagation model and by the assumptions made about the physical scene. The ray tracing software developed for the thesis is created for the sole purpose of understanding RBM. In order to determine the signals interaction with the moving rotor blade in an ideal environment, the ray traced scene is limited to the rotor blades, transmitter and receiver. The signal produced by the simulation is used to derive an algorithm that determines both the azimuth and elevation angle of a transmitter. The derived algorithm is then evaluated on both the deterministic signal to determine its azimuth and elevation accuracy. The algorithm is then tested again on signals in the presence of differing levels of noise to determine the algorithm's performance in a more realistic environment.

>>>>

\section{Thesis Scope}

The scope of this thesis is to investigate whether it is possible to use the effect of RBM to geolocate a transmitter in ideal conditions. Under ideal conditions the received signal is deterministic based upon the propagation model and the initial conditions. This thesis does not intend to include every physical effect but is limited to the effects defined in the propagation model and by the assumptions made about the physical scene. The ray tracing software developed for the thesis is created for the sole purpose of understanding RBM. In order to determine the signals interaction with the moving rotor blade, in an ideal environment, the ray traced scene is limited to the rotor blades, transmitter and receiver. The signal produced by the simulation is used to derive an algorithm that determines both the azimuth and elevation angle of a transmitter. The derived algorithm is then evaluated on the deterministic signal to determine its azimuth and elevation accuracy. The algorithm is then tested again on signals in the presence of differing levels of noise to determine the algorithm's performance in a more realistic environment.

### 2016-11-26 Conclusion.tex
remove red markers, get ready to publish

Over the course of this thesis a three dimensional ray tracer application was built in C++ to produce a RBM signal according to a variety of physical parameters. The physical parameters associated with a helicopter blade and the positioning of transmitter and receiver hardware were swept independently to build knowledge on how they affect the resulting signal. That data was then processed using time frequency analysis to produce a Doppler envelope. An envelope of each parameter step was used to determine an overall trend. Then with those trends an algorithm was derived to find the azimuth and elevation angle of the transmitter with relationship to the helicopter platform. The algorithm accurately finds the azimuth angle and elevation using the data collected in the simulation environment with almost no error in the azimuth estimation and less than 2\% error in elevation angle when blade pitch in 0\textdegree.

>>>>

Over the course of this thesis a three dimensional ray tracer application was built in C++ to produce a RBM signal according to a variety of physical parameters. The physical parameters associated with a helicopter blade and the positioning of transmitter and receiver hardware were swept independently to build knowledge on how they affect the resulting signal. That data was then processed using time frequency analysis to produce a Doppler envelope. An envelope of each parameter step was used to determine an overall trend. Then with those trends an algorithm was derived to find the azimuth and elevation angle of the transmitter with relationship to the helicopter platform. The algorithm accurately finds the azimuth angle and elevation using the data collected in the simulation environment with little error in the azimuth estimation and less than 2\% error in elevation angle when blade pitch in 0\textdegree under a deterministic signal. With the addition of Gaussian noise the algorithm stays relatively constant up to a certain point, then both the azimuth and elevation estimations start to degrade with decreasing elevation angle and increasing noise level. For the simulation scenarios conducted, once the level of noise exceeded 10dB the algorithm started to degrade, but this level will most likely change if other simulation parameters are chosen.

### 2016-11-26 Results_new.tex
remove red markers, get ready to publish

\textcolor{red}{EDITED}

@@

where $\theta_{E_{samp}}$ is the error, in degrees, on either side of the algorithm selected 90\textdegree \space point, $n_{samples}$ is the number of unique sample angles. Therefore in this example there would be 40 unique sample points and a  $\theta_{E_{samp}} = 4.5$\textdegree.

@@

\textcolor{red}{EDITED}

%alt 200
%50 20db noise
%200 12db noise
%400 11.5db noise

%alt 500
%100 20db noise
%400 12db noise
%700 11.5db noise

@@

(text on trends in error with noise level at 11.5dB)

@@

(text on trends in error with noise level at 12dB)

@@

(text on trends in error with noise level at 20dB)

@@

From Table \ref{tab:elevation_error_percent} we can see elevation estimation error is extremely small for the given elevation angles. There is a trend that is evident in the data that as the elevation angle decreases the amount of error is increasing.

\textcolor{red}{EDITED}

% Elevation error table
\begin{table}
\begin{center}
    \begin{tabular}{ | l | l | l | l |}
    \hline
    Altitude & Elevation Angle & Average Elevation Error in (deg) & Average Percent Error \\ \hline
     200 & 76\textdegree &0.5533\textdegree & 0.7284\%  \\ \hline
     200 & 45\textdegree & -0.2647\textdegree & -0.5882\%  \\ \hline 
     200 & 26.5\textdegree & \textdegree & \%  \\ \hline
     500 & 78.7\textdegree & 0.7923\textdegree & 1.0069\% \\ \hline
     500 & 51.34\textdegree & -0.3978\textdegree & -0.7747\%  \\ \hline 
     500 & 35.5\textdegree & -3.9893\textdegree & -11.2255\%  \\ \hline
    \end{tabular}
    \caption{Average Elevation Error and Average Elevation Percent Error of Different Elevation Angles 11.5dB Noise}
    \label{tab:elevation_error_percent_11.5db}
\end{center}
\end{table}

(text on trends in error with noise level at 11.5dB)

% Elevation error table
\begin{table}
\begin{center}
    \begin{tabular}{ | l | l | l | l |}
    \hline
    Altitude & Elevation Angle & Average Elevation Error in (deg) & Average Percent Error \\ \hline
     200 & 76\textdegree & 0.6966\textdegree & 0.9170\%  \\ \hline
     200 & 45\textdegree & -2.3497\textdegree & -5.2215\%  \\ \hline 
     200 & 26.5\textdegree & \textdegree & \%  \\ \hline
     500 & 78.7\textdegree & 0.8916\textdegree & 1.1331\% \\ \hline
     500 & 51.34\textdegree & -4.8613\textdegree & -9.4687\%  \\ \hline 
     500 & 35.5\textdegree & 7.8398\textdegree & 22.0606\%  \\ \hline
    \end{tabular}
    \caption{Average Elevation Error and Average Elevation Percent Error of Different Elevation Angles 12dB Noise}
    \label{tab:elevation_error_percent_12db}
\end{center}
\end{table}

(text on trends in error with noise level at 12dB)

% Elevation error table
\begin{table}
\begin{center}
    \begin{tabular}{ | l | l | l | l |}
    \hline
    Altitude & Elevation Angle & Average Elevation Error in (deg) & Average Percent Error \\ \hline
     200 & 76\textdegree & -9.6248\textdegree & -12.6703\%  \\ \hline
     200 & 45\textdegree & 7.8795\textdegree & 17.5099\%  \\ \hline 
     200 & 26.5\textdegree & \textdegree & \%  \\ \hline
     500 & 78.7\textdegree & 1.3260\textdegree & 1.6851\% \\ \hline
     500 & 51.34\textdegree & 4.5056\textdegree & 8.7760\%  \\ \hline 
     500 & 35.5\textdegree & 20.8592\textdegree & 58.6959\%  \\ \hline
    \end{tabular}
    \caption{Average Elevation Error and Average Elevation Percent Error of Different Elevation Angles 20dB Noise}
    \label{tab:elevation_error_percent_20db}
\end{center}
\end{table}

(text on trends in error with noise level at 20dB)

@@

The error in elevation is caused by a limitation on the estimation. Where the Doppler shift is limited by the maximum amount Doppler that the blade rotation can produce shown in equation \ref{eqn:v_radial} by replacing $v_r$ with $v_rMax$ which is:

@@

@@

If we now consider a pitched blade the above method for estimating azimuth and elevation becomes inaccurate due to the increased Doppler between 90\textdegree \space and 270\textdegree \space and the decreased Doppler on the other side. To correct this error we need to correct both the difference calculation and the correct Doppler data accordingly. The uncorrected data is shown in Figure \ref{fig:diff_pitch_noCorrection} and \ref{fig:dop_pitch_noCorrection} with percent azimuth error of 20\% and percent elevation error of 35\%.

>>>>

@@

where $\theta_{E_{samp}}$ is the error, in degrees, on either side of the selected 90\textdegree \space point, $n_{samples}$ is the number of unique sample angles. Therefore in this example there would be 40 unique sample points and a  $\theta_{E_{samp}} = 4.5$\textdegree.

@@

Adding various levels of Gaussian noise to the received signal degrades the performance of the azimuth estimation. The estimation technique is resilient to noise, showing no change in performance until a certain level of noise is achieved. For the simulated altitudes and elevation angles, for the chosen simulation parameters, there is a noticeable degradation in performance of the estimation at small elevation angles starting at 11.5dB of noise.

@@

Table \ref{tab:az_error_and_percent_11db} shows that the performance has not changed for the higher elevation angles, while the smallest elevation angle azimuth estimations show significant error increase. Increasing the level of noise to 12dB starts to affect the azimuth estimation of the 45\textdegree \space and 51.34\textdegree \space elevation angle simulations shown in Table \ref{tab:az_error_and_percent_12db}. Increasing the noise from 11.5dB to 12dB does not show an increased error in elevation angles of 26.5\textdegree \space and 35.5\textdegree.

@@

In order for additional noise to start affecting the simulated elevation angles of 76\textdegree \space and 78.7\textdegree \space,  the added noise is raised to 20dB. Table \ref{tab:az_error_and_percent_20db} shows the azimuth estimation error for 20dB of added noise, where all the elevation angle simulations show an increased azimuth estimation error where elevation angles of 26.5\textdegree \space and 35.5\textdegree are affected the most.

@@

The azimuth estimation does very well under 10dB of noise, after that point it starts to degrade with decreasing elevation angle as well as increasing level of noise. The reason why estimation error start to show at 10dB of added noise can be traced back to the fact that the Doppler profile is estimated using a power threshold on the short time Fourier transform. This means that the added noise level must increase close to or beyond the threshold to start affecting the Doppler profile and the subsequent azimuth estimation.

@@

From Table \ref{tab:elevation_error_percent} we can see elevation estimation error is extremely small for the given elevation angles. There is a trend that as the elevation angle decreases the amount of error is increasing.

@@

The error in elevation, on the deterministic signal, is caused by a limitation on the estimation. Where the Doppler shift is limited by the maximum amount Doppler that the blade rotation can produce shown in equation \ref{eqn:v_radial} by replacing $v_r$ with $v_rMax$ which is:

@@

%------------------------------
%Noise considerations
%------------------------------

The elevation estimation technique is then tested on the received signals with added noise at the same levels that were previously added for azimuth estimation tests. Starting at 11.5dB of noise, Table \ref{tab:elevation_error_percent_11.5db} shows that the error in estimation is significantly increased at elevation angles of 26.5\textdegree \space and 35.5\textdegree. The other elevation angle estimation errors stay relatively similar to estimation under deterministic signals except for 78.7\textdegree which has increased, but not to the same degree as the elevation angles of 26.5\textdegree \space and 35.5\textdegree.

% Elevation error table
\begin{table}
\begin{center}
    \begin{tabular}{ | l | l | l | l |}
    \hline
    Altitude & Elevation Angle & Average Elevation Error in (deg) & Average Percent Error \\ \hline
     200 & 76\textdegree &0.5533\textdegree & 0.7284\%  \\ \hline
     200 & 45\textdegree & -0.2647\textdegree & -0.5882\%  \\ \hline 
     200 & 26.5\textdegree & 1.9321\textdegree & 7.2731\%  \\ \hline
     500 & 78.7\textdegree & 0.7923\textdegree & 1.0069\% \\ \hline
     500 & 51.34\textdegree & -0.3978\textdegree & -0.7747\%  \\ \hline 
     500 & 35.5\textdegree & -3.9893\textdegree & -11.2255\%  \\ \hline
    \end{tabular}
    \caption{Average Elevation Error and Average Elevation Percent Error of Different Elevation Angles 11.5dB Noise}
    \label{tab:elevation_error_percent_11.5db}
\end{center}
\end{table}

Increasing the noise level to 12dB the error in elevation estimation almost doubles for elevation angles of 26.5\textdegree \space and 35.5\textdegree \space shown in Table \ref{tab:elevation_error_percent_12db}. The increased noise also affects the estimation of elevation angles 45\textdegree \space and 51.34\textdegree, increasing the error from -0.5882\% to -5.2215\% and -0.7747\% to -9.4687\% respectively.

% Elevation error table
\begin{table}
\begin{center}
    \begin{tabular}{ | l | l | l | l |}
    \hline
    Altitude & Elevation Angle & Average Elevation Error in (deg) & Average Percent Error \\ \hline
     200 & 76\textdegree & 0.6966\textdegree & 0.9170\%  \\ \hline
     200 & 45\textdegree & -2.3497\textdegree & -5.2215\%  \\ \hline 
     200 & 26.5\textdegree & 3.9078\textdegree & 14.7102\%  \\ \hline
     500 & 78.7\textdegree & 0.8916\textdegree & 1.1331\% \\ \hline
     500 & 51.34\textdegree & -4.8613\textdegree & -9.4687\%  \\ \hline 
     500 & 35.5\textdegree & 7.8398\textdegree & 22.0606\%  \\ \hline
    \end{tabular}
    \caption{Average Elevation Error and Average Elevation Percent Error of Different Elevation Angles 12dB Noise}
    \label{tab:elevation_error_percent_12db}
\end{center}
\end{table}

At the noise level of 20dB the estimation of elevation angle 76\textdegree \space experiences an error of -12.6703\%. While the elevation angle of 78.7\textdegree \space error increases from 1.1331\% to 1.6851\% shown in Table \ref{tab:elevation_error_percent_20db}.

% Elevation error table
\begin{table}
\begin{center}
    \begin{tabular}{ | l | l | l | l |}
    \hline
    Altitude & Elevation Angle & Average Elevation Error in (deg) & Average Percent Error \\ \hline
     200 & 76\textdegree & -9.6248\textdegree & -12.6703\%  \\ \hline
     200 & 45\textdegree & 7.8795\textdegree & 17.5099\%  \\ \hline 
     200 & 26.5\textdegree & 16.7304\textdegree & 62.9791\%  \\ \hline
     500 & 78.7\textdegree & 1.3260\textdegree & 1.6851\% \\ \hline
     500 & 51.34\textdegree & 4.5056\textdegree & 8.7760\%  \\ \hline 
     500 & 35.5\textdegree & 20.8592\textdegree & 58.6959\%  \\ \hline
    \end{tabular}
    \caption{Average Elevation Error and Average Elevation Percent Error of Different Elevation Angles 20dB Noise}
    \label{tab:elevation_error_percent_20db}
\end{center}
\end{table}

From the Table \ref{tab:elevation_error_percent_12db} to Table \ref{tab:elevation_error_percent_20db} the other estimation errors have now increased as well. The increased errors make elevation angles of 26.5\textdegree \space and 35.5\textdegree \space unresolvable. This means that as the noise level increases the other elevation angles will start to exhibit increasing errors in estimated elevation angle to the point where any elevation angle cannot be accurately estimated.

The elevation angle estimation is more susceptible to noise than the azimuth estimation because it uses the maximum value found in the Doppler profile to find elevation angle from equation \ref{eqn:elevation_angle_only}. Therefor when noise is added the doppler profiles maximum value will fluctuate causing errors in elevation angle estimation.

@@

If we now consider a pitched blade, on the deterministic signal, the derived method for estimating azimuth and elevation becomes inaccurate due to the increased Doppler between 90\textdegree \space and 270\textdegree \space and the decreased Doppler on the other side. To correct this error we need to correct both the difference calculation and the correct Doppler data accordingly. The uncorrected data is shown in Figure \ref{fig:diff_pitch_noCorrection} and \ref{fig:dop_pitch_noCorrection} with percent azimuth error of 20\% and percent elevation error of 35\%.
